---
title: "Bar Island Data Analysis"
author: "Sophie Chivers"
date: "May 4th, 2024"
output:
  html_document: 
  pdf: 
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
#| label: Packages
# install.packages("tidyverse")
library(tidyverse)
library(readr)
library(readxl)
library(knitr)
```

```{r}
#| label: Read in data 
bar_24_raw <- read_csv("data/sand_bar_to_bar_island_usage_survey-2024-04-17_08.05.40.csv")

```

```{r}
#| label: clean-bar_24
bar_24 <- bar_24_raw %>%
  select(observed, Number_of_Adult_Pedestrians, Number_of_Children_Pedestrians,Vehicles_that_drove_on_the_Bar_without_parking_or_picking_up_passengers, Vehicles_that_drove_on_to_the_Bar_and_parked, Vehicles_that_drove_on_the_Bar_for_official_purposes__park_vehicle_headed_to_Bar_Island__Bar_Harbor_Police__or_municipal, Dogs_on_leash, Dogs_off_leash, Number_of_instances_of_dogs_defecating_and_owners_picking_it_up, Number_of_instances_of_dogs_defecating_and_owners_NOT_picking_it_up, `Number_of_private_kayak_groups_(arriving_with_personal_vehicles)`, Number_of_private_kayaks_launched, Number_of_other_boats_launched, Number_of_people_fishing, Number_of_people_birding, Number_of_people_in_the_water, `Number_of_wildlife_disturbances_(if_a_dog_disturbs_a_group_of_birds_count_as_one)_`, Number_of_bicyclists, Current_Weather) 

colnames(bar_24) <- c("observed", "adult_pedestrians", "child_pedestrians","vehicles_no_parking", "vehicles_parking", "vehicles_official", "dogs_leash", "dogs_off_leash", "dogs_defecating_pick_up", "dogs_defecating_no_pick_up", "kayak_group", "private_kayaks", "other_boats", "fishers", "birders", "swimmers_waders", "wildlife_disturbances", "bicyclists", "weather")

bar_24 <- bar_24 %>% 
  mutate_if(is.numeric, replace_na, replace = 0) %>%
  filter(!row_number() %in% c(1, 16))

bar_24$observed <- as_datetime(bar_24$observed)

bar_24 <- bar_24 %>%
  mutate(weather = as_factor(bar_24$weather))

bar_24_with_total <- bar_24 %>%
  mutate(total_pedestrians = rowSums(bar_24[ , c("adult_pedestrians", "child_pedestrians")], na.rm=TRUE))

# bar_24 <-mutate(weather = fct_relevel(bar_24$weather, c("clear", "partly cloudy", "overcast", "rain")))

```

```{r}
#| label: pivot-bar_24-longer

bar_24_long <- bar_24 %>%
  pivot_longer(! c(observed, weather),
    names_to = "use",
    values_to = "number") %>%
  group_by(use) %>%
  mutate(general_use = case_when(
         use == "adult_pedestrians" ~ "pedestrians", 
         use == "child_pedestrians" ~ "pedestrians", 
         use == "vehicles_no_parking" ~ "vehicles", 
         use == "vehicles_parking" ~ "vehicles",
         use == "vehicles_official" ~ "vehicles",
         use == "dogs_leash" ~ "dogs", 
         use == "dogs_off_leash" ~ "dogs",
         use == "dogs_defecating_pick_up" ~ "dog_poo", 
         use == "dogs_defecating_no_pick_up" ~ "dog_poo",
         use == "kayak_group" ~ "kayak", 
         use == "private_kayaks" ~ "kayak", 
         use == "other_boats" ~ "boats", 
         use == "fishers" ~ "other_rec", 
         use == "birders" ~ "other_rec", 
         use == "swimmers_waders" ~ "other_rec", 
         use == "wildlife_disturbances" ~ "wildlife_disturbances", 
         use == "bicyclists" ~ "bikes")) %>%
  mutate(date = date(observed)) %>%
  mutate(month = month(date)) %>%
  filter(!month == "1") 
```

```{r}
#| label: Create daily summary df

bar_24_long_daily_summary <- bar_24_long %>% 
  group_by(date, general_use) %>%
  summarize(n = sum(number))
```

# Visualization of Bar Island Data

## Time Series Plots

```{r}
#| label: Line plots
bar_24_long_daily_summary %>%
  filter(!general_use == "dog_poo") %>%
  group_by(general_use) %>%
  ggplot(
    aes( 
      x = date, 
      y = n, 
      color = general_use)
  ) + 
  geom_line() + 
  labs(
    title = "Seasonality of Use in 2023", 
    x = "Date", 
    y = "Number of Users"
  )

# bar_24 %>%
#   ggplot(
#     mapping = aes(
#       x = observed,
#       y = adult_pedestrians)) +
#     geom_line() +
#   labs(
#     title = "Adult Pedestrians on the Bar Over Time")
#
# bar_24 %>%
#   ggplot(
#     mapping = aes(
#       x = observed, 
#       y = child_pedestrians)) +
#     geom_line() + 
#   labs(
#     title = "Number of Children Walking on Bar over Time"
#   )
# 
# # bar_24 %>%
# #   ggplot(
# #     mapping = aes(
# #       x = observed, 
# #       y = total_pedestrians)) +
# #     geom_line() 
# 
# bar_24_long %>%
#   select(c(date, general_use, number)) %>%
#   group_by(general_use) %>%
#   ggplot(
#     aes(
#       x = date,
#       y = number,
#       color = general_use, na.rm)) +
#   geom_line() +
#   labs(
#     title = "Frequency of Use over Time"
#   )
# 
# # bar_24 %>% 
# #   select(adult_pedestrians, child_pedestrians, observed) %>%
# #   ggplot( 
# #     aes(x=observed, y=n, group=name, color=name)) +
# #     geom_line()
```

## Bar Plots

```{r}
#| label: Bar plots

# bar_24_long %>%
#   filter(general_use == c("bikes", "boats", "dogs", "kayak", "other_rec", "pedestrians", "vehicles", "wildlife_disturbances")) %>%
#   ggplot(
#     aes(
#       fill = general_use, 
#       x = general_use, 
#       y = number)) + 
#   geom_bar(position= "dodge", stat = "identity") + 
#   labs( 
#     title = "Bar Island Visitation by Type of Use")

bar_24_long %>%
  group_by(general_use) %>%
  ggplot( 
    aes(
      fill = general_use,
      x = weather, 
      y = number)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs( 
    title = "2023 Bar Island Bar Useage by Type of Use and Weather", 
    x = "Weather",
    y = "Number of Visitors", 
    ) 

bar_24_long %>%
  filter(general_use == c("pedestrians", "vehicles", "kayak", "bikes", "dogs")) %>%
  ggplot( 
    aes(
      fill = general_use,
      x = general_use, 
      y = number
    )) +
  geom_bar(stat = "identity") + 
  labs(
    title = "Total Number of Visitors in 2023 by Use", 
    x = "Use Type", 
    y = "Numbe of Visitors"
  )

```

## Comparisons with 2006 Data

The 2006 data was taken with slightly different methodologies than the 2023 data. The 2006 data was collected during all daylight hours, regardless of tidal height. The 2023 data was only taken in shifts surrounding the low tide. The 2023 shifts covered the entirety of bar exposure during daylight hours. These data should be mostly comparable, but it is important to note that 2006 data may have collected relatively higher counts per day than than the 2023 data.

The 2006 data was taken over 15 days.

The 2023 data was taken over 39 days.

```{r}
#| label: Effort calculations

d <- bar_24 %>%
  mutate(date = date(observed)) %>%
  distinct(date) %>%
  nrow()

# specific_use_per_day <- bar_24_long %>% 
#   group_by(use) %>%
#   summarize(n_per_day = sum(number)/d) %>%
#   print()
# 
# specific_use <- bar_24_long %>% 
#   group_by(use) %>%
#   summarize(no = sum(number)) %>%
#   print()

# specific_use_tibble <- full_join(specific_use, specific_use_per_day)

use_per_day_23 <- bar_24_long %>% 
  filter(general_use != "dog_poo") %>%
  group_by(general_use) %>% 
  summarize(n_per_day = sum(number)/d) %>%
  mutate(year = as.numeric("2023"))
  
use_23 <- bar_24_long %>% 
  filter(general_use != "dog_poo") %>%
  group_by(general_use) %>%
  summarize(n = sum(number)) %>%
  mutate(year = as.numeric("2023"))

data_2006 <- read_excel("data/2006 data.xlsx") %>%
  mutate(general_use = case_when(
    general_use == "cars" ~ "vehicles", 
    general_use == "bikes" ~ "bikes", 
    general_use == "pedestrians" ~ "pedestrians",
    general_use == "kayak" ~ "kayak", 
    general_use == "dogs" ~ "dogs"))

use_per_day_06 <- data_2006 %>%
  select(c(general_use, n_per_day, year))

use_06 <- data_2006 %>%
  select(c(general_use, n, year))

use_per_day <- bind_rows(use_per_day_06, use_per_day_23) %>%
  mutate(year = as.character(year)) 

use <- bind_rows(use_06, use_23) %>%
  mutate(year = as.character(year))

# Things I want to improve here: 
# 1) Make one column for 2023 and one for 2006 "n per days"
# 2) Round to the nearest visitor 

kable(use, 
      caption = "Table of Total Counted Visitors by Use Type for 2006 and 2023")

kable(use_per_day, 
      caption = "Table of Average Users per day by Use Type for 2006 and 2023")

```

```{r}
#| label: Comparative plots

use_per_day %>% 
  filter(!general_use %in% c("boats", "other_rec", "wildlife_disturbances")) %>%
  ggplot( 
    aes(
      x = general_use, 
      y = n_per_day, 
      fill = year
    )) + 
  geom_bar(position = "dodge", stat = "identity") + 
  labs(
    title = "Average Visitation (per day) in 2006 and 2023"
  )

use %>% 
  filter(!general_use %in% c("boats", "other_rec", "wildlife_disturbances")) %>%
  ggplot( 
    aes(
      x = general_use, 
      y = n, 
      fill = year
    )) + 
  geom_bar(position = "dodge", stat = "identity") + 
  labs(
    title = "Total Recorded Visitation in 2006 and 2023"
  )

```
